---
title: "Reproducible Research - Course Project 1"
author: "Ryan Maxfield"
date: "June 8, 2018"
output: html_document
---


Data
====

Dataset: https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip

Dataset variables:

        .       steps - Number of steps taken in 5-minute intervals
        
        .       date - Date the measurement was taked YYYY-MM-DD format
        
        .       interval - Identifies the 5-minute interval

Dataset is stored as .csv and contains a total of 17,568 observations


##1. Code for reading in the dataset and/or processing the data

Obtain and load the dataset

```{r}

if(!file.exists("repdata%2FData%2Factivity.zip")) {
        temp<-tempfile()
        download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)
        unzip(temp)
        unlink(temp)
}

dx<-read.csv("activity.csv")
dz<-dx
```


##2. Histogram of the total number of steps taken each day

Aggregate data and obtain the total number of steps per day and graph using ggplot2

```{r}
da<-aggregate(steps~date,dx,sum)
library(ggplot2)
g<-ggplot(da,aes(steps))
g+geom_histogram(binwidth = 500,color="black",fill="red")+
        xlab("Number of Steps/Day")+
        ggtitle("Histogram Total Number of Steps Taken Each Day",subtitle="Reproducible Research - Course Project 1")

```


##3. Mean and median number of steps taken each day

Mean and median number of steps per day available using in summary

```{r}
summary(da$steps)
```


##4. Time series plot of the average number of steps taken

Aggregate data and obtain the average number of steps per 5-minute period and graph using ggplot2

```{r}
db<-aggregate(steps~interval,dx,mean,na.rm=TRUE)
h<-ggplot(db,aes(interval,steps))
h+geom_line(color="darkred")+
        xlab("Five Minute Interval")+
        ylab("Number of Steps")+
        ggtitle("Average Number of Steps per Five Minute Interval")
```


##5. The 5-minute interval that, on average, contains the maximum number of steps

Determine the 5-minute interval with the average maximum number of steps

```{r}
db[which.max(db$steps),1]
```


##6. Code to describe and show strategy for imputing missing data

Determine number of missing values
```{r}
sum(is.na(dx$steps))
```

Where no data is available replace with the calculated average (mean) using tapply

```{r}
dx$date<-as.Date(dx$date,"%m/%d/%Y")
index<-which(is.na(dx$steps))
l<-length(index)
avg<-with(dx,tapply(steps,date,mean,na.rm=TRUE))
dy<-mean(avg,na.rm=TRUE)
for(i in 1:l){
        dx[index[i],1]<-dy
}
```


##7. Histogram of the total number of steps taken each day after missing values are imputed

Aggregate data and obtain the total number of steps per day and graph using ggplot2

```{r}
dc<-aggregate(steps~date,dz,sum)
library(ggplot2)
g<-ggplot(dc,aes(steps))
g+geom_histogram(binwidth = 500,color="black",fill="red")+
        xlab("Number of Steps/Day")+
        ggtitle("Histogram Total Number of Steps Taken Each Day (Imputed)",subtitle="Reproducible Research - Course Project 1")
```


##8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

Use xyplot function from lattice library

```{r}
dx$date <- as.Date(strptime(dx$date, format="%Y-%m-%d"))
dx$day <- weekdays(dx$date)
for (i in 1:nrow(dx)) {
    if (dx[i,]$day %in% c("Saturday","Sunday")) {
        dx[i,]$day<-"weekend"
    }
    else{
        dx[i,]$day<-"weekday"
    }
}
dwk<-aggregate(dx$steps~dx$interval+dx$day,dx,mean)
names(dwk)<-c("interval","day","steps")
library(lattice)
xyplot(steps~interval|day,dwk,type="l",layout=c(1, 2), 
    xlab="Interval",ylab="Number of Steps")
```

##The END


